---
title: "dplyr-tutorial"
author: "Joan Rojas"
date: "24 Mar 2015"
runtime: shiny
output: 
  html_document:
    theme: flatly
---



This is my first rmd document. Recently I completed the dplyr tutorial from Hadley Wickham. You can find Part I [here](http://datascience.la/hadley-wickhams-dplyr-tutorial-at-user-2014-part-1/) and Part II [here](http://www.r-bloggers.com/hadley-wickhams-dplyr-tutorial-at-user-2014-part-2/).

First I load the libraries and the dataset used in this tutorial. In this case, the `flights`dataset is explored. It contains information of outgoing flights from the Houston airport. The variables of this dataset are the following:

```{r message = FALSE, echo = FALSE}
library(ggplot2)
library(dplyr)
flights <- tbl_df(read.csv("flights.csv", stringsAsFactors = FALSE))
flights$date <- as.Date(flights$date)
names(flights)
```

The dplyr package has five main verbs:

* `filter` used to select rows based on some criteria
* `select` used to select columns based on some criteria
* `mutate` to create or modify variables
* `arrange` to arrange the rows
* `summarise`

It is very interesting to combine them using the opertaor `%>%`which can be interpreted as **then**.

The figure below shows the average distance flight the selected carrier flew in that year.

```{r echo = FALSE, fig.align='center' , fig.height= 2 , fig.width= 2, fig.star = TRUE }
c <- flights %>% distinct(carrier) %>% select(carrier)
selectInput("carrier", label = "Select Carrier", choices = c$carrier,
            multiple = TRUE)

renderPlot({
flights %>% select(carrier, dist) %>%
  filter(carrier %in% as.character(input$carrier)) %>%
  group_by(carrier) %>%
  summarise(distance = mean(dist, na.rm = TRUE)) %>%
  ggplot(aes( y = distance , x = carrier)) + geom_bar(stat = "identity")
})
```
